clusters

### 2015-07-13 cluster.rmd
clustering

trial<- (sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)})) %>% as.data.frame
colnames(trial)[names(trial) %in% '.']<-'var'
ggplot(trial, aes_string(x='var')) + geom_histogram(aes(y=..density..))
skewness(trial$var)
kurtosis(trial$var)
mean(trial$var)
sd(trial$var)
trial$scale.var<- scale(trial$var)
ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
# ggplot still looks bimodal after scaling

@@

clust$suc<- sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)}) %>% scale()

## should we make return.pdol bimodal??
math.cbrt <- function(x) {
        sign(x) * abs(x)^(1/3)
}
qqPlot(all7c$return.pdol %>% math.cbrt,
       main='success')
trial<- (all7c$return.pdol %>% math.cbrt) %>% as.data.frame
colnames(trial)[names(trial) %in% '.']<-'var'
ggplot(trial, aes_string(x='var')) + geom_histogram(aes(y=..density..))
skewness(trial$var)
kurtosis(trial$var)
trial$scale.var<- scale(trial$var)
ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
#take cube root of return.pdol and scale it to clust

clust$return.pdol<- all7c$return.pdol %>% math.cbrt %>% scale()

@@

clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
                                      suc, hrs.mlsto), 2, 15, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)

@@

# 11 clusters should be good

# have a look at dendrogram with 11 clusters

x<- dist(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 
         method = 'euclidean')
fit<- hclust(x, method = 'ward.D2')
plot(fit)

rect.hclust(fit, k=11, border='red')

@@

now want 11 clusters, based on k means optimisation, have a look at average characteristics

@@

k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 11)

@@

all7c$groups<- k1$cluster
kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, return.pdol, suc, groups)
all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
all7c.plot$groups<- rownames(all7c.plot)
k2.melt<- melt(all7c.plot)

#summary means of all numeric variables in 5 clusters
clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
bal<- sum(all7c$balance.mlsto)
hrs<-  sum(all7c$hrs.mlsto)
clust.summary$pc.jobs<- round( tab[,2]*100, 1)
clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
        select(2)
clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
        select(2)
clust.summary$job.count<- table(all7c$groups) %>% as.data.frame %>% select(Freq)

@@

# add medians
all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))

@@

clust.summary$bal<- ddply(all7c, .(groups), summarise, bal= round(sum(balance.mlsto), 1)) %>% 
        select(2)

@@

>>>>

trial<- (sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)})) %>% as.data.frame
colnames(trial)[names(trial) %in% '.']<-'var'
ggplot(trial, aes_string(x='var')) + geom_histogram(aes(y=..density..))
skewness(trial$var)
kurtosis(trial$var)
mean(trial$var)
sd(trial$var)
trial$scale.var<- trial$var/sd(trial$var)
ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
# ggplot still looks bimodal after scaling

@@

clust$suc<- sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)})
clust$suc<- clust$suc/sd(clust$suc)

## should we make return.pdol bimodal??
math.cbrt <- function(x) {
        sign(x) * abs(x)^(1/3)
}
qqPlot(all7c$return.pdol %>% math.cbrt,
       main='success')
trial<- (all7c$return.pdol %>% math.cbrt) %>% as.data.frame
colnames(trial)[names(trial) %in% '.']<-'var'
ggplot(trial, aes_string(x='var')) + geom_histogram(aes(y=..density..))
skewness(trial$var)
kurtosis(trial$var)
trial$scale.var<- trial$var/sd(trial$var)
ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
#take cube root of return.pdol and scale it to clust

clust$return.pdol<- math.cbrt(all7c$return.pdol)
clust$return.pdol<- all7c$return.pdol/sd(all7c$return.pdol)

## should we make balance.mlsto bimodal??

qqPlot(all7c$balance.mlsto,
       main='balance')
qqPlot(sapply(all7c$balance.mlsto, function(x){sign(x)*log(abs(x) + 1)}),
       main='balance')
trial<- (sapply(all7c$balance.mlsto, function(x){sign(x)*log(abs(x) + 1)})) %>% as.data.frame
colnames(trial)[names(trial) %in% '.']<-'var'
ggplot(trial, aes_string(x='var')) + geom_histogram(aes(y=..density..))
skewness(trial$var)
kurtosis(trial$var)
trial$scale.var<- trial$var/sd(trial$var)
ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
#take cube root of return.pdol and scale it to clust

clust$balance.mlsto<- sapply(all7c$balance.mlsto, function(x){sign(x)*log(abs(x) + 1)})
clust$balance.mlsto<- clust$balance.mlsto/sd(clust$balance.mlsto)

@@

clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto), 2, 30, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)

@@

# 16 clusters should be good

# have a look at dendrogram with 16 clusters

x<- dist(clust %>% select(inv.mlsto, balance.mlsto), 
         method = 'euclidean')
fit<- hclust(x, method = 'ward.D2')
plot(fit)

rect.hclust(fit, k=16, border='red')

@@

now want 16 clusters, based on k means optimisation, have a look at average characteristics

@@

k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto), 16)

@@

all7c$group.num<- k1$cluster
kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, return.pdol, suc, group.num)
all7c.plot<- ddply(kmeans.plot, .(group.num), numcolwise(mean))
all7c.plot$group.num<- rownames(all7c.plot)
k2.melt<- melt(all7c.plot)

#summary means of all numeric variables in 5 clusters
clust.summary<- all7c.plot %>% select(-group.num) %>% round(2)
tab<- table(all7c$group.num) %>% prop.table %>% as.data.frame
bal<- sum(all7c$balance.mlsto)
hrs<-  sum(all7c$hrs.mlsto)
clust.summary$pc.jobs<- round( tab[,2]*100, 1)
clust.summary$bal.pc<- ddply(all7c, .(group.num), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
        select(2)
clust.summary$hrs.pc<- ddply(all7c, .(group.num), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
        select(2)
clust.summary$job.count<- table(all7c$group.num) %>% as.data.frame %>% select(Freq)

@@

# add medians
all7c.med<- ddply(kmeans.plot, .(group.num), numcolwise(median)) %>% round(2)

@@

clust.summary$bal<- ddply(all7c, .(group.num), summarise, bal= round(sum(balance.mlsto), 1)) %>% 
        select(2)

#investigate clusters
all7c[all7c$group.num == 12,] %>% select(inv.mlsto, suc, pc.majpos, majority.pos,
                                         code.client, client.meaninv, code.director, code.ProjEng,
                                         return.pdol, balance.mlsto, Year, timespan, pc.pro,
                                         pc.director, JD.Second, Business, client.count, Discipline,
                                         code.client, Billing.Type) %>% summary

@@

Visualise confidence intervals in terms of 'success'

```{r CI, echo=FALSE}
library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
# principal component analysis

#add clust columns to all7c
all7c$clust.inv<- clust$inv.mlsto
all7c$clust.bal<- clust$balance.mlsto
all7c$clust.cost<- clust$cost.mlsto
all7c$clust.return<- clust$return.pdol
all7c$clust.hours<- clust$hrs.mlsto
all7c$clust.suc<- clust$suc

#convert cluster groups to integer

all7c$group.num<- as.integer(all7c$group.num)
p<- prcomp(~clust.inv + clust.bal+ clust.cost, data=all7c, centre=T)
# summary(p)
biplot(p, xlabs = as.character(all7c$group.num))
#putlers in-house biplot shows centroids
bpCent(p, clsAsgn = all7c$group.num, centroids=TRUE, xlabs = as.character(all7c$Business))

##try ggbiplot
library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
#visualise clusters
q<- ggbiplot(p, obs.scale = 1, var.scale = 1, 
              groups = all7c$group.num %>% as.factor, ellipse = TRUE,
             labels = all7c$Business,
              circle = TRUE)
q <- q + scale_color_discrete(name = '')
q <- q + theme(legend.direction = 'vertical', 
               legend.position = 'right')
q

#visualise factors within clusters
g<- ggbiplot(p, obs.scale = 1, var.scale = 1, 
              groups = all7c$Business %>% as.factor(), ellipse = TRUE, 
              circle = TRUE)
g <- g + scale_color_discrete(name = '')
g <- g + theme(legend.direction = 'vertical', 
               legend.position = 'right')
g

#confusion matrix of groups vs. categorical variables

#Business
table(all7c$group.num, all7c$Business)
#notes
# only single NFP - combine to institution
all7c[all7c$Business %in% 'NFP',]$Business<- 'institution'
#combine person and friend
all7c[all7c$Business %in% 'friend',]$Business<- 'person'
#fabricator labelled manufactorer/supplier
all7c[all7c$Business %in% 'fabricator',]$Business<- 'manufacturer/supplier'
all7c[all7c$Business %in% 'membrane fabricator',]$Business<- 'manufacturer/supplier'
#combine remedial building services to building services
all7c[all7c$Business %in% 'remedial building service',]$Business<- 'building services'
#combine gov and council
all7c[all7c$Business %in% 'council',]$Business<- 'gov'
all7c$Business<- droplevels(all7c$Business)

View counts of each category in each group as a heatmap

Order groups by profitability and size
